Audio Interface of an Active Perception Navigation Aid for People with Visual Impairments

### 2019-07-11 paper.tex
Added experiment description, system implementation

\begin{abstract}
	%Our aim is to build a portable active navigation system for people with visual impairments that uses active perception techniques and a combination of feedback modes to identify a scene and guide a user to their destination.
	%Such a system requires a non-visual feedback interface and in this paper we investigate the effectiveness of a spatial audio tone with a varying pitch component, played through bone-conducting headphones, in conveying the pan and tilt angles of a target to the user in a pointing task.
	%The secondary goal is to determine how variations to the pitch's rate of change affect the user's performance.
	%For this, we conducted a set of experiments with both blindfolded participants and participants with visual impairments and found that, with some limitations, a spatialised audio tone with a varying pitch component can successfully convey a target's pan and tilt angles. 
\keywords{Human-machine interface \and vision impairment \and spatialised sound \and varying pitch, pointing task}
\end{abstract}

@@

\begin{figure}
  \centering
  \subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.4\columnwidth]{figures/tango_headphone.png}}
~
  \subfloat[]{\label{fig:participant}\includegraphics[width=0.25\columnwidth]{figures/vi_participant.pdf}}
  \caption{Pictures of the Tango device and bone-conducting headphones the system is based on (left) and both in use in an experiment (right).}\label{fig:tango-headphone}
\end{figure}

@@

%Our ultimate goal is to deliver a navigation system that can guide a user with visual impairments on the last leg of their journey (e.g.\ to a specific aisle and shelf in a shop) using only a mobile phone.
%A large amount of data need to be translated from a visual form into a format that is useful to people with limited or no vision.
%We therefore plan to use a combination of voice, audio and vibration cues to translate visual navigation data as effectively as possible and overcome the bandwidth limitation of the human ear. 
%However, for this paper, we only considered the spatialised audio mode and its variation in pitch in order to determine its effectiveness in conveying a target's pan and tilt angles.

%The system is based on a concept proposed in~\cite{bellotto2013,lock2017portable}, which uses a Google Tango device.
%The navigation system is based on a Google Tango tablet, an Android-based device that comes equipped with an RGB-D camera to sense colour and depth and combines an inertial measurement unit (IMU) with powerful and robust landmark recognition and image processing algorithms to localise itself in real-time.
%An added benefit of this platform is its familiar, compact form-factor which will help overcome the hurdle of user-acceptance and usability.
%The guidance instructions are presented through a set of bone-conducting headphones (see \cref{fig:tango-headphone}) that are placed externally on a user's head so that the system does not occlude the perception of real-world sounds and does not interfere with a user's normal hearing function.

%A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}, where the arrows indicate the direction of the flow of information.
%When the user taps the Tango's screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the device's current position and orientation.
%The audio generator then produces a tone based on the difference between the device and target's positions and sends it to the audio output channel, which plays it back to the user.
%The WiFi recording module is constantly monitoring the different parameter values of the device and target's positions, as well as the system's output, and records it to a remotely stored datafile.

%\begin{figure}
  %\centering
  %\includegraphics[clip=true, trim=0 120 80 50, width=1.0\columnwidth]{figures/pipeline.pdf}
  %\caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }\label{fig:pipeline}
%\end{figure}

@@

The guidance target is presented to the user in terms of pan and elevation angles, indicating the angle adjustments that are required to point the device camera at the target location, as shown in \cref{fig:cam-coords}.
Spatialised audio signals are well-suited to the task, displaying similar levels of performance to vocal feedback, but with less cognitive load and higher resolution\cite{klatzky2006cognitive}.
However, the audio guidance signals are delivered to the user with a set of bone-conducting headphones that bypass the user's outer-ear structure and since this structure plays an important role in localising a sound source's elevation\cite{blauert1969sound}, an alternative mode of conveying the elevation angle is required.
For this work, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle. 
The pan angle can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conducting headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.\todo{cite bone cnducting work}

@@

@@

\subsection{Apparatus}

\subsection{Procedure}

@@

%\section{Portable Navigation System}\label{sec:portable-navigation}

%\subsection{System Setup}

%Our ultimate goal is to deliver a navigation system that can guide a user with visual impairments on the last leg of their journey (e.g.\ to a specific aisle and shelf in a shop) using only a mobile phone.
%A large amount of data need to be translated from a visual form into a format that is useful to people with limited or no vision.
%We therefore plan to use a combination of voice, audio and vibration cues to translate visual navigation data as effectively as possible and overcome the bandwidth limitation of the human ear. 
%However, for this paper, we only considered the spatialised audio mode and its variation in pitch in order to determine its effectiveness in conveying a target's pan and tilt angles.

%The system is based on a concept proposed in~\cite{bellotto2013,lock2017portable}, which uses a Google Tango device.
%The navigation system is based on a Google Tango tablet, an Android-based device that comes equipped with an RGB-D camera to sense colour and depth and combines an inertial measurement unit (IMU) with powerful and robust landmark recognition and image processing algorithms to localise itself in real-time.
%An added benefit of this platform is its familiar, compact form-factor which will help overcome the hurdle of user-acceptance and usability.
%The guidance instructions are presented through a set of bone-conducting headphones (see \cref{fig:tango-headphone}) that are placed externally on a user's head so that the system does not occlude the perception of real-world sounds and does not interfere with a user's normal hearing function.

%A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}, where the arrows indicate the direction of the flow of information.
%When the user taps the Tango's screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the device's current position and orientation.
%The audio generator then produces a tone based on the difference between the device and target's positions and sends it to the audio output channel, which plays it back to the user.
%The WiFi recording module is constantly monitoring the different parameter values of the device and target's positions, as well as the system's output, and records it to a remotely stored datafile.

%\begin{figure}
  %\centering
  %\includegraphics[clip=true, trim=0 120 80 50, width=1.0\columnwidth]{figures/pipeline.pdf}
  %\caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }\label{fig:pipeline}
%\end{figure}

%\subsection{Audio Interface}

%The audio component is responsible for conveying the 2D position of a target on the vertical plane, shown by \cref{fig:cam-coords}, in terms of pan and tilt angles.
%The audio signal is a continuous \SI{10}{\hertz} sinusoidal sound wave that is played to the user through bone-conducting headphones.
%We use a sinusoidal wave here because it is relatively simple to manipulate and analyse and will be replaced by a richer, more natural tone at a later stage.

%The audio is spatialised in the pan dimension using the OpenAL\footnote{\url{http://openal.org}} sound library's HRTF, while the tilt angle is conveyed by varying the audio tone's pitch.
%We use this approach because the external set of bone-conducting headphones plays the sound through a user's cheekbones instead of their outer ears, bypassing the ears' pennae that provide the ability to localise elevated sound sources~\cite{roffler1968factors,algazi2001elevation}.
%This makes an HRTF less effective in conveying the target's elevation and therefore needs to be conveyed using another method; by varying the tone's pitch in this case.
%The difference between the target's angular position and the device's angular orientation are used to generate the audio signals.

%\subsubsection{Pan Direction}

%The pan angle ($\phi$) describes the angle which the user needs to rotate the camera vector, $C$, around the $z$-axis shown in \cref{fig:cam-coords}, i.e.\ how far the target is to the left or right of the user.
%We use an HRTF to add a spatial element to the audio tone that is played to the user, placing the sound source at the target's location. 
%Humans determine a sound source's pan direction by subconsciously analysing the inter-aural time difference (IID) between a sound reaching both of your ears: the greater the IID, the greater the perceived angular distance to the sound source~\cite{wightman1992dominant}.

%We implement OpenAL's default HRTF, based on the MIT's KEMAR dataset\footnote{\url{http://sound.media.mit.edu/resources/KEMAR.html}}, to generate a sinusoidal sound wave based on the difference between the user and target's positions.
%The library takes position values as input and outputs a tone based on the angle between the two position vectors.

%\subsubsection{Tilt Direction}

%The system adjusts the tone's pitch ($f$) to convey the target's tilt angle ($\theta$) w.r.t.\ $C$. 
%Here, a high pitch means the target is above $C$ (i.e.\ the user should look up) and a low pitch means the target is below $C$ (the user should look down).
%This high/low association scheme is chosen because humans naturally associate high-pitched sounds with higher objects and lower-pitched noises with lower objects~\cite{pratt1930spatial,blauert1997spatial}.
%We use a logarithmic, octave-based gain function for the pitch, since an increase in octave provides a distinct perceptible change while keeping the timbre roughly constant~\cite{shepard1964circularity}.

%We wish to determine how the gradient, $m$, of the pitch gain as a function of $\theta$ (i.e.\ $m = \frac{df}{d\theta}$) affects a user's performance, e.g.\ does a greater $m$ lead to an increased target acquisition rate?
%For this we select 3 different pitch settings, referred to henceforth as $m_l$, $m_m$ and $m_h$ for the smallest, middle and highest gain gradients respectively. 
%To find these gradients, we set the maximum and minimum limits for $\theta$ and the maximum and minimum frequencies for $f$ for each respective setting.
%Furthermore, for the sake of consistency and between the different settings, each gradient generates the same pitch when $C$ is level with the target's tilt.

%After practical tests with the device and headphones, we set the neutral, on-target tone to $f_{512} =$ \SI{512}{\hertz} for its audibility.
%For $m_l$, we set $f_{max}$ and $f_{min}$ to 1 octave higher and lower than $f_{512}$, giving limits of \SI{1024}{\hertz} and \SI{256}{\hertz} respectively.
%For $m_m$, the $f_{max}$ and $f_{min}$ are set to 2 octaves higher and lower than $f_{512}$ (\SI{2048}{\hertz} and \SI{128}{\hertz}) and $m_h$ to 3 octaves higher and lower (\SI{4096}{\hertz} and \SI{64}{\hertz}).
%These limits were selected for practical reasons, because the bone conducting headphones have low volume gain at the frequency spectrum extremes, making it difficult to hear. 
%The values of $m_l, m_m, m_h$ are visualised in \cref{fig:pitch-gain}.

%\begin{figure}
  %\centering
  %\includegraphics[clip, trim=0 0 0 40, width=1.0\columnwidth]{figures/pitch_gain_functions.png}
  %\caption{The pitch gain functions for the 3 different gain settings ($y$-axis in log scale). }\label{fig:pitch-gain}
%\end{figure}

%We set the range of angles the system can convey to $[-\frac{\pi}{2}; \frac{\pi}{2}]$; anything outside this range implies that the target is behind the user and was not considered in this series of experiments. 
%The \SI{0}{\radian} position is directly in front of the user.
%With these limits, $f$ can be determined with

%\begin{equation}\label{eq:frequency}
  %f(\theta) = 2^{b\theta + c},
%\end{equation}
%where $b$ and $c$ for each setting can be calculated according to the given setting's pitch and angle limits.% and a linear graph equation (i.e.\ $f - f_1 = b(\theta - \theta_1), b=\frac{\Delta f}{\Delta\theta}$).

%\section{Target Search Experiment}\label{sec:experiments}

%We conducted a set of experiments to evaluate the audio interface's effectiveness at driving a user to point the camera at a target in the pan and tilt dimensions.
%Furthermore, we wish to see how changes to $m$ affect a user's performance.

%We recruited 2 groups of participants for the experiments.
%Group $G1$ consists of 42 young adults (mostly students) with healthy eyesight, with 10 males and 32 females being $20\pm1.5$ years old on average.
%This group was blindfolded throughout the experiment and forms the bulk of the participant population.
%Group $G2$ is made up of 3 legally blind adults with little to no light perception (3 males, average age of $40\pm8$ years). 
%This group makes up less than 10\% of the total experiment population, but provided valuable subjective feedback that will help improve the system before evaluating it with a larger group of people with vision impairments. 
%All participants joined the experiment on a voluntary basis with written consent and have no other impairments that affected their interpretation of the audio guidance cues.

%For the experiment, the participants were given a Tango tablet device, running an app written for this experiment, and a pair of AfterShockz Sportz3' bone conducting headphones (both pictured in use in \cref{fig:participant}) and were blindfolded where appropriate.
%The app generates a virtual target, one at a time at a constant \SI{2}{\metre} from the participant, and uses the device's pose measurements to determine $\phi$ and $\theta$.
%These angles are then used to generate and present the user with the audio signal to direct them to the target. 
%As the participant rotates the camera, $\phi$ and $\theta$ are adjusted and the audio signal changes to reflect these changes.
%When a participant believes that they are pointing at the target, they tap the screen which generates a new target target which they are then directed to. 
%Each participant was given 28 targets in total that were randomly generated and equally spread across 4 quadrants to avoid clustering. 
%After the 28 targets were found, $m$ was adjusted and the experiment was repeated. 
%The order in which the gradient setting was changed for each participant was randomised to minimise the learning effect.
%The app streams various data on the current device and target poses to a laptop computer via a WiFi connection in real-time. 
%Before the experiment session started, each participant was given some time to familiarise themselves with the device and was given the opportunity to learn what the $f_{512}$ `on-target' and on-centre tone sounds like that will mean that they are pointing at the target.

%The measure of performance here is the accuracy with which a user points the camera towards a target.
%The accuracy is taken as the angular difference between $C$ and the target's position at the time the participant tapped the screen.
%A smaller error indicates more accurate results.
%The pan and tilt results are separated to understand the effect of changing the gradient setting has on either dimension's accuracy.
%To minimise the speed/accuracy biases in the results, we asked the participants to focus on finding the targets as well as they could without worrying about the time it took.

>>>>

\begin{abstract}
	%Our aim is to build a portable active navigation system for people with visual impairments that uses active perception techniques and a combination of feedback modes to identify a scene and guide a user to their destination.
	%Such a system requires a non-visual feedback interface and in this paper we investigate the effectiveness of a spatial audio tone with a varying pitch component, played through bone-conducting headphones, in conveying the pan and tilt angles of a target to the user in a pointing task.
	%The secondary goal is to determine how variations to the pitch's rate of change affect the user's performance.
	%For this, we conducted a set of experiments with both blindfolded participants and participants with visual impairments and found that, with some limitations, a spatialised audio tone with a varying pitch component can successfully convey a target's pan and tilt angles. 
  The ActiVis project is attempting to build a mobile guidance aid to help people with limited vision search for objects within an unknown environment.
  This system requires an effective non-visual interface to transmit guidance instructions to the user through means of bone-conducting headphones.
  To this end we propose a audio-based interface that uses a spatialised signal to convey a target's position on the horizontal plane. 
  The position on the median plan is given by adjusting the tone's pitch to overcome audio localisation issues that arise with bone-conducting headphones. 
  This interface is validated through a set of experiments. 
\keywords{Human-machine interface \and vision impairment \and spatialised sound \and varying pitch, pointing task}
\end{abstract}

@@

\begin{figure}[t]
  \centering
  \subfloat[]{\label{fig:tango-headphone}\includegraphics[width=0.4\columnwidth]{figures/tango_headphone.png}}
~
  \subfloat[]{\label{fig:participant}\includegraphics[width=0.25\columnwidth]{figures/vi_participant.pdf}}
  \caption{Pictures of the Tango device and bone-conducting headphones the system is based on (left) and both in use in an experiment (right).}\label{fig:tango-headphone}
\end{figure}

@@

@@

The guidance target is presented to the user in terms of pan and elevation angles, indicating the angle adjustments that are required to point the device camera at the target location, as shown in \cref{fig:cam-coords}.
Spatialised audio signals are well-suited to the task, displaying similar levels of performance to vocal feedback, but with less cognitive load and higher resolution\cite{klatzky2006cognitive}.
However, the audio guidance signals are delivered to the user with a set of bone-conducting headphones that bypass the user's outer-ear structure and since this structure plays an important role in localising a sound source's elevation\cite{blauert1969sound}, an alternative mode of conveying the elevation angle is required.
For this work, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle. 
The pan angle can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conducting headphones\cite{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.

@@

A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}, where the arrows indicate the direction of the flow of information.
When the user taps the Tango's screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the device's current position and orientation.
The audio generator then produces a tone based on the difference between the device and target's positions and sends it to the audio output channel, which plays it back to the user.
The WiFi recording module is constantly monitoring the different parameter values of the device and target's positions, as well as the system's output, and records it to a remotely stored datafile.

\begin{figure}
  \centering
  \includegraphics[clip=true, trim=0 120 80 50, width=0.8\columnwidth]{figures/pipeline.pdf}
  \caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }\label{fig:pipeline}
\end{figure}

@@

To test the interface's effectiveness at directing a user towards a target area, a set of experiments were conducted with a simple pointing task. 
This pointing task is set up to capture the difference between the targets' actual position and the positions the participants' perceived them to be.

For the experiments, the participants are given a Tango tablet device running an app specifically written for these experiments. 
This app generates a set of virtual targets and presents them to the participants one at a time. 
The targets are generated at a constant \SI{2}{\meter} from the participant and their the pan and elevation angles are uniformly generated across 4 quadrants to avoid clustering.
Each target's angular position is communicated to the participant through the audio interface, the output of which is adjusted in real-time as the participant points the device around. 
When a participant was confident that they were on-target, i.e.\ hearing the audio front-on at \SI{512}{\hertz}, they tapped the screen, marking the location and generating the next target.
The targets' positions are all set relative to the Tango device's coordinate system which is tracked using the Tango localisation API.\
A total of 28 targets were generated per participant.

2 groups of participants were recruited for the experiments. 
Group \textit{G1} consisted of 10 young adults with normal eyesight that were blindfolded for the experiments, and group \textit{G2} contains 2 people with severe vision impairments. 
Both groups were given some time with the system before the experiments commenced to familiarise themselves with the system, the audio signal's behaviour and the \SI{512}{\hertz} on-level tone. 
To minimise the speed/accuracy biases in the results, we asked the participants to focus on finding the targets as well as they could without worrying about the time it took.

@@
